\documentclass[a4paper,11pt]{book}
\usepackage{parskip}
\usepackage[top=1.5cm, left=1cm, right=1cm, bottom=2.5cm]{geometry}
\usepackage{fontspec,xunicode}
\setmainfont[Scale=.95]{Trebuchet MS}
\begin{document}
\chapter{Introduction}
\section{System Architecture}
\subsection{Application}
Android will ship with a set of core applications. All applications are written
using the Java programming language. They call the APIs provided by the Application Framework.
\subsection{Application Framework}
Developer have full access to the same framework APIs used by the core
applications. The application architecture is designed to simplify the reuse of
componets; any application can publish its capabilities and any other
application may then make use of those capabilities.

The widgets provided by Application Framework:
\begin{description}
\item[Activity Manger]Manage the lifecycle of activities.
\item[WindowMangaer] Manage the Window
\item[Content Provider] Allowing application to access data of other application
or share their data
\item[View System] To build application, including List, Gird, Textbox, Button
and graphics and WebView.
\item[Notification Manager] Showing information on status bar.
\item[Package Manager] Manage the application installed.
\item[Resource Manager]
\item[Hardware Services]
\end{description}
\section{Android Initialization Procedure}
\begin{enumerate}
  \item Android will boot linux kernel's Init first
  \item Startup the linux daemons, initializing:
    \begin{itemize}
      \item usbd, to manage usb connection
      \item adbd, Android Debug Bridge, to manage ADB connection. 
      \item debuggered, manage debug requests
      \item rild, Manage wireless connections
    \end{itemize}
  \item Start a Zygote thread, this thread will:
    \begin{itemize}
      \item Initialize a Dalvik instance.
      \item Load a Class and a Listener for Socket requests
      \item Create a Dalvik instance for managing applicaiton threads
    \end{itemize}
  \item Initialize runtime thread. It will:
  \begin{itemize}
   \item Initialize ServiceManager 
  \item register ServieManager. It will be the default Context manager for Binder
  service
  \end{itemize}
  \item After initialization of runtime, runtime thread wil send a request to
  Zygote to start system service, and Zygote will create an Dalvik VM instance for
  system service, and start the system server.
  \item System Server will start the native system services, including
  \emph{Surface Flinger} and \emph{Audio Flinger}. All those services will
  reigster to Service Manager as the target of IPC.
  \item System Service will start the Android Manage Service, all Android
  Services will register to the Service Manager.
  \item After loading all service, system will waiting for the running of
  applications. Every application will start a new thread. 
\end{enumerate}
\chapter{Android Kernel Architecture}
\section{The relation between Linux and Android}
\section{Android is not Linux}
Android:
\begin{itemize}
\item It doesn't have Local window system, it doesn't need to use the X server
\item It doesn't have glibc support, it uses Bionic Libc instead
\item It is not a standard Linux kernel. For example the Gold-Fish platform
based on ARM, and yaffs2FLASH filesystem.
\item Drivers. Android enhanced the Linux drivers. Mainly on:
  \begin{description}
  \item[Android Binder] Based on OpenBinder framework, provide InterProcess
  Communication. The source code located in
  \verb|drivers/staging/android/binder.c|
  \item[Power Manager (PM)] A Powermanager opitimized for Embeded Device. The
  source code is under \verb|kernel/power|
  \item[Low Memory Killer] More flexible OOM(Out Of Memory) Scheme, the source
  code is in drivers/staing/android/lowmemorykiller.c
  \item[Ashmem] Share memory between threads, and provide a scheme for Kernel to
  manage and recycle the memory. Source code in mm/ashmem.c
  \item[Android PMEM(Physical)] Provide a continuous physical memory. In
  \verb|drivers/misc/pmem.c|
  \item[Android Logger]
  \item[Android Alam] A timer to wake the device up from sleeping. At
  \verb|drivers/rtc/alarm.c|
  \item[USB Gadget Driver] in \verb|drivers/usb/gadget|
  \itme[Android Ram Console] Provide debugging functions. Allowing debug log
  to write into a RAM Console device. It's a Buffer based on RAM. in
  \verb|drivers/staging/android/ram_console.c|
  \item[Android timed device] 
  \item[Yaffs2 Filesystem] in \verb|fs/yaffs2/| For flash storage (Yet another
  flash file system).
  
  \end{description}
\end{itemize}

\end{document}

